---
title: "Happiness_Project"
author: "Emma Horton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}
library(dplyr)
```


Read in Dataset

```{r}
getwd()
setwd("/Users/emmahorton/Documents/GitHub/world_happiness_report/Datasets/")
df <- read.csv("WorldHappinessReport.csv")
#View(df)

```

View Head
```{r}
#head(df)
#tail(df)
```

View Missing
```{r}
summary(df)
```
Cleaning Data for Regional Indicator

```{r}
df[df$Regional.Indicator == "",]
```

Missing Regions Replacement
```{r}
cleaning <- df %>%
  mutate(Regional.Indicator = case_when(
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Angola' ~ "Southern Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Belize' ~ "Central America",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Bhutan' ~ "South Asia",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Central African Republic' ~ "Central Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Congo (Kinshasa)' ~ "Central Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Cuba' ~ "Caribbean",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Czechia' ~ "Central Europe",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Djibouti' ~ "East Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Eswatini' ~ "Southern Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Guyana' ~ "Anglophone Caribbean",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Oman' ~ "Middle East",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Qatar' ~ "Middle East",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Somalia' ~ "East Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Somaliland region' ~ "East Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'South Sudan' ~ "North Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Sudan' ~ "North Africa",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Suriname' ~ "South America",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Syria' ~ "Middle East",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Trinidad and Tobago' ~ "Caribbean",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'Turkiye' ~ "Central and Eastern Europe",
    (is.na(Regional.Indicator) | Regional.Indicator == "") & Country.Name == 'State of Palestine' ~ "West Asia",
    TRUE ~ Regional.Indicator
  ))

```

Verify Region does not contain empty values
```{r}
cleaning[cleaning$Regional.Indicator == "", ]
```
```{r}
View(cleaning)
```


Cleaning Data Median Values for Numeric Variables

```{r}
medians <- cleaning %>% 
  summarize(across(where(is.numeric), ~median(., na.rm = TRUE)))

cleaning <- cleaning %>%
  mutate(across(where(is.numeric), ~ifelse(is.na(.), medians[[cur_column()]], .)))

summary(cleaning)
```
```{r}
unique(cleaning$Regional.Indicator)
```
Attach Continents
```{r}
cleaning <- df %>%
  mutate(Continent = case_when(
    Regional.Indicator %in% c("South Asia", "Commonwealth of Independent States", "Southeast Asia", 
                  "East Asia", "Middle East", "West Asia") ~ "Asia",
    Regional.Indicator %in% c("Central and Eastern Europe", "Western Europe", "Central Europe") ~ "Europe",
    Regional.Indicator %in% c("Middle East and North Africa", "Southern Africa", "Sub-Saharan Africa", 
                  "Central Africa", "East Africa", "North Africa") ~ "Africa",
    Regional.Indicator %in% c("Latin America and Caribbean", "North America and ANZ", "Central America", 
                  "Caribbean", "Anglophone Caribbean", "South America") ~ "Americas",
    Regional.Indicator == "North America and ANZ" ~ "Oceania",
    TRUE ~ "Other" 
  ))
```
Outlier Detection on Life Ladder

```{r}
boxplot(Life.Ladder ~ Continent, data = cleaning,
        main = "Life Ladder Distribution by Continent",
        xlab = "Continent", ylab = "Life Ladder")

```



Outlier Detection on Log.GDP.Per.Capita
```{r}
boxplot(Log.GDP.Per.Capita ~ Continent, data = cleaning,
        main = "Log.GDP.Per.Capita Distribution by Continent",
        xlab = "Continent", ylab = "Log.GDP.Per.Capita")
```

Outlier Detection on Social Support
```{r}
boxplot(Social.Support ~ Continent, data = cleaning,
        main = "Social Support Distribution by Continent",
        xlab = "Continent", ylab = "Social Support")
```
Outlier Detection on Healthy Life Expectancy

```{r}
boxplot(Healthy.Life.Expectancy.At.Birth ~ Continent, data = cleaning,
        main = "Healthy Life Expectancy Distribution by Continent",
        xlab = "Continent", ylab = "Healthy Life Expectancy")
```


Outlier Detection on Freedom.To.Make.Life.Choices

```{r}
boxplot(Freedom.To.Make.Life.Choices ~ Continent, data = cleaning,
        main = "Freedom in Life Choice Distribution by Continent",
        xlab = "Continent", ylab = "Freedom in Life Choice")
```

Outlier Detection on Generosity


```{r}
boxplot(Generosity ~ Continent, data = cleaning,
        main = "Generosity Distribution by Continent",
        xlab = "Continent", ylab = "Generosity")
```


Outlier Detection on Corruption Perception
```{r}
summary(cleaning)
boxplot(Perceptions.Of.Corruption ~ Continent, data = cleaning,
        main = "Perception of Corruption Distribution by Continent",
        xlab = "Continent", ylab = "Perception of Corruption")
```


Outlier Detection on Positive Affect by Region
```{r}
boxplot(Positive.Affect ~ Continent, data = cleaning,
        main = "Positive.Affect Distribution by Continent",
        xlab = "Continent", ylab = "Positive.Affect")
```


Outlier Detection on Negative Effect
```{r}
boxplot(Negative.Affect ~ Continent, data = cleaning,
        main = "Negative Effect Distribution by Continent",
        xlab = "Continent", ylab = "Negative Effect")
```

Save DataFrame in a Pretty Name
```{r}
cleaning -> happydf
```

Export Data to CSV
```{r}
write.csv(happydf,"/Users/emmahorton/Documents/GitHub/world_happiness_report/Cleaning/HappyDataFrame", row.names=FALSE)
```


